---
title: "Lab5"
author: "Max Snellenberger"
date: "2025-05-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Lab 2 Task
```{r}
library(tidyverse)
library(sf)
library(tmap)
library(leaflet)
library(dplyr)
library(spdep)
library(ggplot2)

# counties <- sf::read_sf("../data/CBW/County_Boundaries.shp")
# counties %>% sf::st_make_valid()

#load datasets
counties <- sf::read_sf("../data/CBW/County_Boundaries.shp") %>% 
  sf::st_make_valid()
dams <- sf::read_sf("../data/CBW/Dam_or_Other_Blockage_Removed_2012_2017.shp") %>% 
  sf::st_make_valid()
streams <- sf::read_sf("../data/CBW/Streams_Opened_by_Dam_Removal_2012_2017.shp") %>% 
  sf::st_make_valid()

# aspatial
bmps <- read_csv("../data/CBW/BMPreport2016_landbmps.csv")
```

```{r}
#add a column for the combined state and county fips in counties
counties <- counties %>% mutate(., FullFIPS = paste0(STATEFP10, COUNTYFP10))

#create a column for trimmed fips in bmps
bmps <-  bmps %>% mutate(., FIPS.trimmed = substr(GeographyName, 1, 5))

#add bmps within each county
summed_bmps <- bmps %>%
  group_by(FIPS.trimmed) %>%
  summarise(AmountCredited = sum(AmountCredited, na.rm = TRUE))
summed_bmps

#join both datasets by fips
CountyBMP <- right_join(counties, summed_bmps, by = c("FullFIPS" = "FIPS.trimmed" ))
glimpse(CountyBMP)
#remove any values that are NA
CountyBMP <- CountyBMP %>%
  filter(!is.na(AmountCredited))

# Remove rows with missing geometries
CountyBMP <- CountyBMP[!is.na(st_geometry(CountyBMP)), ]

# #make the map
# tm_shape(CountyBMP) + 
#   tm_polygons(fill = "AmountCredited")

CountyBMP

palette <- colorBin(palette = "YlGnBu", 
                    domain = CountyBMP$AmountCredited,
                    bins = 5)

leaflet(data = CountyBMP) %>%
  addProviderTiles("CartoDB.Positron") %>%  #Add basemap
  addPolygons(
    fillColor = ~palette(AmountCredited),
    fillOpacity = .9,
    color = "white",
    weight = 1,
    label = ~paste0(
  NAME10, ": $", formatC(AmountCredited, format = "f", digits = 0, big.mark = ",")
) %>% lapply(htmltools::HTML)
) %>%
  addLegend(pal = palette, values = ~AmountCredited
  )
```

Lab 3
```{r}
NEprojected <- read_sf("./NortheastStates.shp")

# Ensure NEprojected is in sf format
NEprojected_sf <- st_as_sf(NEprojected)

# Transform to WGS84 (lat/lon) for use with Leaflet
NEprojected <- st_transform(NEprojected_sf, crs = 4326)

# Ensure NEprojected is in sf format
NEprojected_sf <- st_as_sf(NEprojected)

# Check the class and structure of the sf object
print(class(NEprojected_sf))  # Should be "sf"
str(NEprojected_sf)

NEnb <- spdep::poly2nb(NEprojected, queen = TRUE)

pts <- st_centroid(NEprojected)

IDWlw <- nb2listwdist(NEnb, as(pts, "Spatial"), style = "W", zero.policy = TRUE)
IDWlw$weights[1]

# Ensure the lisa_quadrant column exists and is correct
table(NEprojected_sf$lisa_quadrant)

# Define a color palette for the LISA quadrants
color_palette <- colorFactor(
  palette = c("red", "blue", "yellow", "green"), 
  domain = NEprojected_sf$lisa_quadrant
)

# Calculate Local Moran's I for each observation
local_moran <- localmoran(NEprojected$B01001e10, IDWlw)

# Add the Local Moran's I results to your dataset
NEprojected$local_moran_i <- local_moran[, 1]
NEprojected$p_value <- local_moran[, 5]
mean_observed <- mean(NEprojected$B01001e10)
# Create LISA quadrant classification
NEprojected$lisa_quadrant <- ifelse(NEprojected$local_moran_i > 0 & NEprojected$B01001e10 > mean_observed, "High-High",
                                    ifelse(NEprojected$local_moran_i > 0 & NEprojected$B01001e10 < mean_observed, "Low-Low", 
                                           ifelse(NEprojected$local_moran_i < 0 & NEprojected$B01001e10 > mean_observed, "High-Low", 
                                                  "Low-High")))


# Create the interactive leaflet map
leaflet(data = NEprojected_sf) %>%
  addProviderTiles("CartoDB.Positron", group = "CartoDB Positron") %>%
  addTiles(group = "OpenStreetMap") %>%
  addProviderTiles("Esri.WorldImagery", group = "Esri Satellite") %>%

  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(
    fillColor = ~color_palette(lisa_quadrant),  # Map colors to LISA quadrants
    color = "black",  # Border color
    weight = 1,  # Border weight
    opacity = 1,  # Border opacity
    fillOpacity = 0.7,  # Fill opacity
    popup = ~paste0("<strong>LISA Quadrant:</strong> ", lisa_quadrant,
                    "<br><strong>p-value:</strong> ", round(p_value, 4)),
     highlight = highlightOptions(
      weight = 3,
      color = "white",
      fillOpacity = 0.7,
      bringToFront = TRUE
    )
  ) %>%
  addLegend(
    position = "bottomright",
    pal = color_palette,
    values = NEprojected_sf$lisa_quadrant,
    title = "LISA Quadrants",
    opacity = 1
  ) %>%
  addLayersControl(
    baseGroups = c("CartoDB Positron", "OpenStreetMap", "Esri Satellite"),,
    options = layersControlOptions(collapsed = FALSE)
  )


```

```{r}
state <- read_sf("./static_mapping/oh_counties.gpkg")
labels <- read_sf("./static_mapping/oh_places.gpkg")
parks <-  read_sf("./static_mapping/oh_parks.gpkg")
rivers <- read_sf("./static_mapping/oh_rivers.gpkg")

local <- read_sf("./static_mapping/oh_counties.gpkg") %>%
  dplyr::filter(NAME %in% c("Portage", "Summit"))
oh_region <-  st_bbox(local) %>%
  st_as_sfc()

local.map <- tm_shape(state, bbox = oh_region) + 
  tm_borders() +
  tm_shape(labels) + 
  tm_fill(fill_alpha = 0.5) + 
  tm_labels("NAME", size = 0.5) +
  tm_shape(parks) + 
  tm_polygons(fill = "FEATTYPE", fill_alpha = .5, palette = "brewer.greens") +    
  tm_shape(rivers) + tm_lines(col = "blue")
  
local.map
```
```{r}
# Reproject all layers to WGS84
state_wgs <- st_transform(state, 4326)
labels_wgs <- st_transform(labels, 4326)
parks_wgs <- st_transform(parks, 4326)
rivers_wgs <- st_transform(rivers, 4326)
local_wgs <-  st_transform(local, 4326)

#crop
state_crop <- st_intersection(state_wgs, local_wgs)
labels_crop <- st_intersection(labels_wgs, local_wgs)
parks_crop <- st_intersection(parks_wgs, local_wgs)
rivers_crop <- st_intersection(rivers_wgs, local_wgs)

# #define centroids for labels
#labels_centroids <- st_centroid(labels_crop)

# Create a color palette for parks by FEATTYPE
feattypes <- unique(parks$FEATTYPE)
park_pal <- colorFactor("Greens", domain = feattypes)

# Leaflet map
leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  
  # Counties borders
  addPolygons(
    data = local_wgs,
    fill = FALSE,
    color = "black",
    weight = 1,
    group = "County Borders"
  ) %>%
  
  # Parks with fill by FEATTYPE
  addPolygons(
    data = parks_crop,
    fillColor = ~park_pal(FEATTYPE),
    fillOpacity = 0.5,
    color = "darkgreen",
    weight = 0.5,
    group = "Parks",
    popup = ~paste("Park Type:", FEATTYPE)
  ) %>%
  
  # Rivers
  addPolylines(
    data = rivers_crop,
    color = "blue",
    weight = 1,
    group = "Rivers"
  ) %>%
#   
#   # Place labels
#   addLabelOnlyMarkers(
#   data = labels_crop,
#   label = ~NAME,
#   labelOptions = labelOptions(noHide = TRUE, textsize = "10px"),
#   group = "Place Labels"
# )
  
  # # Add Layers Control
  # addLayersControl(
  #   overlayGroups = c("County Borders", "Parks", "Rivers", "Place Labels"),
  #   options = layersControlOptions(collapsed = FALSE)
  # ) %>%
  # 
  # # Zoom to bounding box
  # fitBounds(
  #   lng1 = st_bbox(oh_region)$xmin,
  #   lat1 = st_bbox(oh_region)$ymin,
  #   lng2 = st_bbox(oh_region)$xmax,
  #   lat2 = st_bbox(oh_region)$ymax
  # )

```
Questions
1. I thought the labs this semester were very challenging but helpful to learn how to use R and to learn more about spatial analysis. I learned how to break projects down into steps to make them more achievable when the problem is too large to tackle all at once. I liked how the labs were challenging problems, and I wish that I had been able to put more time and effor into them this semester.
